Reasoning Graphically about Quantum Computation

### 2008-02-24 papers/reasoning-graphically/paper.tex
more pictures
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1359 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

To model the behaviour of quantum systems these generators must
satisfy certain equations.

Each family $(\delta, \epsilon, \alpha)$ is governed by a set of
equations defining a forms a \emph{classical structure}
\cite{Coecke2006Quantum-Measure,Coecke2006POVMs-and-Naima} with a
\emph{phase group}.  The requisite equations state that $(\delta,
\epsilon )$ form a cocommutative comonoid:
\begin{gather}
  (\delta \otimes \id{Q}) \circ \delta 
   =   ( \id{Q} \otimes \delta) \circ \delta  
  \tag{A}\label{comonoid-assoc}
  \\
  (\epsilon \otimes \id{Q}) \circ \delta 
  = (\id{Q} \otimes \epsilon)\circ \delta 
  = \id{Q} \tag{U} \label{comonoid-unit}
  \\
  \delta = \sigma \circ \delta. \tag{C}\label{comonoid-comm}
\end{gather}
Their adjoints $(\delta^\dag,\epsilon^\dag)$ form a commutative
monoid':
\begin{gather}
  \delta^\dag \circ (\delta^\dag \otimes \id{Q})
  =    \delta^\dag   \circ ( \id{Q} \otimes \delta^\dag) 
  \tag{A}\label{monoid-assoc}
  \\
  \delta^\dag  \circ (\epsilon^\dag \otimes \id{Q}) 
  = \delta^\dag \circ (\id{Q} \otimes \epsilon^\dag)
  = \id{Q} \tag{U} \label{monoid-unit}
  \\
  \delta^\dag = \delta^\dag \circ
  \sigma. \tag{C}\label{monoid-comm}.
\end{gather}
Futher, these jointly satisfy the \emph{isometry} and \emph{frobenius} laws:
\begin{gather}
  \delta^\dag \circ \delta = \id{Q} \tag{I}\label{isometry}
  \\
  (\delta^\dag \otimes \id{Q}) \circ (\id{Q}\otimes \delta)
  = \delta \circ \delta^\dag. \tag{F}\label{frobenius}
\end{gather}

>>>>

To model the behaviour of quantum systems certain additional equations
must be satisfied.  Each family $(\delta, \epsilon)$ must form a
\emph{classical structure}
\cite{Coecke2006Quantum-Measure,Coecke2006POVMs-and-Naima}:  the pair
$(\delta, \epsilon)$  should form a cocommuative comonoid, the pair
$(\delta^\dag,\epsilon^\dag)$ must form a commutative monoid, and
together they must satisfy the isometry and frobenius equations. In
addition the family $\alpha$ must form an abelian group with $\alpha^\dag
= -\alpha$.  These equations are presented graphically below for the
green family;  the red vertices obey all the same equations
\begin{description}
\item[Comonoid Laws] 
\[
\begin{array}{ccccccccccccccc}
  \inlinegraphic{3.5em}{comonoid-assoc1}  &=\;&
  \inlinegraphic{3.5em}{comonoid-assoc2}
  &\qquad\qquad&
  \inlinegraphic{3.7em}{comonoid-unit1}  &=\;&
  \inlinegraphic{3.7em}{comonoid-unit2}  &=\;&
  \inlinegraphic{3.7em}{comonoid-unit3} 
  &\qquad\qquad&
  \inlinegraphic{3.5em}{comonoid-comm1}  &=\;&
  \inlinegraphic{3.5em}{comonoid-comm2}
\end{array}
\]
\item[Monoid Laws] 
\[
\begin{array}{ccccccccccccccc}
  \inlinegraphic{3.5em}{monoid-assoc1}  &=\;&
  \inlinegraphic{3.5em}{monoid-assoc2}
  &\qquad\qquad&
  \inlinegraphic{3.7em}{monoid-unit1}  &=\;&
  \inlinegraphic{3.7em}{monoid-unit2}  &=\;&
  \inlinegraphic{3.7em}{monoid-unit3} 
  &\qquad\qquad&
  \inlinegraphic{3.5em}{monoid-comm1}  &=\;&
  \inlinegraphic{3.5em}{monoid-comm2}
\end{array}
\]
\item[Isometry] 
\item[Frobenius]
  \[
  \begin{array}{ccccc}
    \inlinegraphic{3.5em}{frobenius1}  &=\;&
    \inlinegraphic{3.5em}{frobenius2} &=\;&
    \inlinegraphic{3.5em}{frobenius3}
  \end{array}
  \]
  \item[Abelian Unitary Group]
\end{description}

Each family $(\delta, \epsilon, \alpha)$ is governed by a set of
equations defining a forms a \emph{classical structure}
 with a
\emph{phase group}.  The requisite equations state that $(\delta,
\epsilon )$ form a cocommutative comonoid:
\begin{gather}
  (\delta \otimes \id{Q}) \circ \delta 
   =   ( \id{Q} \otimes \delta) \circ \delta  
  \tag{A}\label{comonoid-assoc}
  \\
  (\epsilon \otimes \id{Q}) \circ \delta 
  = (\id{Q} \otimes \epsilon)\circ \delta 
  = \id{Q} \tag{U} \label{comonoid-unit}
  \\
  \delta = \sigma \circ \delta. \tag{C}\label{comonoid-comm}
\end{gather}
Their adjoints $(\delta^\dag,\epsilon^\dag)$ form a commutative
monoid':
\begin{gather}
  \delta^\dag \circ (\delta^\dag \otimes \id{Q})
  =    \delta^\dag   \circ ( \id{Q} \otimes \delta^\dag) 
  \tag{A}\label{monoid-assoc}
  \\
  \delta^\dag  \circ (\epsilon^\dag \otimes \id{Q}) 
  = \delta^\dag \circ (\id{Q} \otimes \epsilon^\dag)
  = \id{Q} \tag{U} \label{monoid-unit}
  \\
  \delta^\dag = \delta^\dag \circ
  \sigma. \tag{C}\label{monoid-comm}.
\end{gather}
Futher, these jointly satisfy the \emph{isometry} and \emph{frobenius} laws:
\begin{gather}
  \delta^\dag \circ \delta = \id{Q} \tag{I}\label{isometry}
  \\
  (\delta^\dag \otimes \id{Q}) \circ (\id{Q}\otimes \delta)
  = \delta \circ \delta^\dag. \tag{F}\label{frobenius}
\end{gather}
